---
title: "features.csv"
output: github_document
author: Peter Hickey 
date: "`r Sys.Date()`"
---

**CellRanger** requires a CSV file describing the antibody derived tag (ADT) and hashtag oligo (HTO) features, which we refer to as `features.csv`.
The `features.csv` file is the concatenation of a `ADT_features.csv` and a `HTO_features.csv` file.

This document briefly describes how we generated the `features.csv` file.

## `ADT_features.csv`

We used the [TotalSeq™-A Mouse Universal Cocktail, V1.0](https://www.biolegend.com/ja-jp/products/totalseq-a-mouse-universal-cocktail-v1-21733) from BioLegend.
All antibodies are anti-mouse unless otherwise specified.
The associated metadata for this panel is available in the Excel file [`TotalSeq_A_Mouse_Universal_Cocktail_v1_128_Antibodies_199901_Barcodes.xlsx`](TotalSeq_A_Mouse_Universal_Cocktail_v1_128_Antibodies_199901_Barcodes.xlsx) (source: [BioLegend](https://www.biolegend.com/Files/Images/BioLegend/totalseq/TotalSeq_A_Mouse_Universal_Cocktail_v1_128_Antibodies_199901_Barcodes.xlsx)):

```{r}
library(here)
library(readxl)
x <- read_excel(
  here(
    "data", 
    "sample_sheets",
    "TotalSeq_A_Mouse_Universal_Cocktail_v1_128_Antibodies_199901_Barcodes.xlsx"))
knitr::kable(x)
```

I have used the information in `TotalSeq_A_Mouse_Universal_Cocktail_v1_128_Antibodies_199901_Barcodes.xlsx` to create a [`ADT_features.csv`](ADT_features.csv) file that is compatible with **CellRanger**.

```{r}
y <- data.frame(
  id = x$DNA_ID,
  # NOTE: Tidy up to remove most "anti-mouse" (and variations thereon) prefixes
  #       and replace ",", with ";" to avoid parsing issues of the resultant 
  #       CSV.
  name = gsub(",", ";", gsub("anti-mouse ", "", x$Description)),
  read = rep("R2", nrow(x)),
  pattern = rep("^(BC)", nrow(x)),
  sequence = x$Sequence,
  feature_type = rep("Antibody Capture", nrow(x)))
write.csv(
  y,
  file = here("data", "sample_sheets", "ADT_features.csv"),
  row.names = FALSE,
  quote = TRUE)
knitr::kable(y)
```

## `HTO_features.csv`

There are 27 pools of samples (although not all were progresed to scRNA-seq) and each has a unique HTO label
Specifically, we used combinatorial HTO labelling whereby each pool is labelled with k HTOs (here k=2):

```{r}
sample_metadata_df <- read_excel(
  here("data/sample_sheets/G000304_Broomfield.sample_metadata.xlsx"))
knitr::kable(sample_metadata_df)
```

Based on this, I constructed a `HTO_features.csv` file for use with **CellRanger**.

```{r}
z <- read.csv(here("data/sample_sheets/HTO_features.csv"))
# NOTE: `feature_type` must be 'Antibody Capture' rather than 
#       'Multiplexing Capture' because we're not actually using CellRanger's 
#       demultiplexing feature.
z$feature_type <- "Antibody Capture"
knitr::kable(z)
```

## `features.csv`

We concatenate `ADT_features.csv` and `HTO_features.csv` to generate `features.csv`.

```{r}
features <- rbind(y, z)
write.csv(
  features,
  file = here("data", "sample_sheets", "features.csv"),
  row.names = FALSE,
  quote = TRUE)
```
